
---
title: "xlr"
development: 
  mode: release
output: github_document
---


```{r, include = FALSE}
# ![](vignettes/www/xlr.svg)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
<br>

> A set of tools to help data scientists work interactively between 
> __R__ and __spreadsheets__. 

<hr>

<!-- badges: start -->
`r badger::badge_devel("eauleaf/xlr", "blue")`
`r badger::badge_last_commit("eauleaf/xlr")`
<!-- badges: end -->


<br>


## Setup
``` r
# install.packages("remotes")
remotes::install_github('eauleaf/xlr', build_vignettes = TRUE)
library(xlr)
```

## Example

```{r example, eval = FALSE}
# Send R data to workbook
iris |> split(f = iris$Species) |> xl()
```

## Testing

Because several `xlr` functions require interactive use, reporting issues would be helpful. 

## Contribute

`xlr` has taken hundreds of hours to write, test, and re-write. I hope it makes your work easier. If you do find it helpful, please consider paying it forward by helping Ukraine: https://u24.gov.ua/


