---
title: "xlr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{xlr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(xlr)
```

## 

# Overview

`xlr` is simple; it's fewer than 10 functions designed to make it easier to 
process data in R and view that data in spreadsheets. 

Spreadsheets excel at presenting 2-D data in a human-friendly formats, and 
spreadsheets are the workhorse of the business community. Unfortunately, 
spreadsheets don't have the flexibility or horsepower of R, so I want to 
use R for the heavy lifting, and then pump those tables out to spreadsheets 
to review or share the data. 

If you work in R, and you work in Libreoffice or Excel, then you'll probably find
`xlr` helpful. 


# Getting Started

To install `xlr`, run:

```{r install-xlr, eval = F}
remotes::install_github('eauleaf/crudr')
```

`xlr` is brand-new, so please provide feedback. Try to break the functions. If 
you can, tell me. Once the package is decently stable, I'll submit it to CRAN. 


## `view_xl()` 
The Most Important `xlr` Function

The inspiration for `view_xl()` is dplyr's `view()` function. `view()` is
awesome since it's kind of a stripped-down spreadsheet view. I can look at the
data, then change my code, view the data again, change my code again, etc.

Sometimes, however, I need to review a lot of data from a lot of dataframes. So,
I need dplyr's `view` function, but implemented as a workbook in Libreoffice or
Excel. This is what `view_xl` is for. **`view_xl()`** is an opinionated way to
review dataframes via spreadsheet, implemented with a lot of hidden and
automated conveniences.


### Use
To use `view_xl()`, call a dataframe or a list of dataframes, like so:
`view_xl(iris)`
`view_xl(iris, mtcars)`
`list(flower_data = iris, car_data = mtcars) |> view_xl()`
`iris |> split(f = iris$Species) |> view_xl()`

Your spreadsheet is temporary. Close it and it's gone. If you want to save 
it, then just specify an input path, like: 


The path argument uses `here::here()`, so you can sspecify ... our default directory with 


Install package `clipr` and 
Check with `clipr::dr_clipr()`
