---
title: "xlr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{xlr}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

```

<!-- ![](www/xlr.svg) -->

```{r setup, include = FALSE}
library(xlr)
```


## ***R + Spreadsheets***


<br>
<br>

The package is somewhere between *dev and beta.*

--- **Testing and feedback would be helpful** ---



<br>

## Overview:

Spreadsheets are a great tool for presenting 2-dimensional data in a
reader-friendly format and are still the workhorse of the business
community, but spreadsheets don't have the data processing power of R,
so its common for data professionals to work in both spreadsheets and R.
However, working in both tools can be frustrating. It's time expensive
to read data back forth between the two, and most of the time, I just
want a series of decent-looking tables pushed out to a workbook. **xlr** is
an effort to reduce the friction that comes from needing to work in both
spreadsheets and R.

<br>

## Functions:

-   **`xl()`** -- quickly output a list of datasets as spreadsheets in a
    workbook
-   **`enlist()`** -- make named lists without additional nesting
-   **`copy_for_xl()`** -- copy an R dataset to clipboard memory,
    formatted for pasting to a spreadsheet
-   **`paste_from_xl()`** -- copy spreadsheet data to clipboard memory,
    formatted for pasting to R
-   **`enscript()`** -- produce the code to re-create an R object
-   **`repeated()`** -- flag all repeated observations in a vector
-   **`sys_open()`** -- request the OS to open files, folders, or URLs
-   **`list_iron()`** -- flatten lists of embedded lists to a named, single-file list
-   **`entibble()`** -- create tibbles with default behavior useful for
    spreadsheets

<br>


## Details and Examples:

```{r download-xlr, eval = F}
# Install
remotes::install_github('eauleaf/xlr', build_vignettes = TRUE)
library(xlr)
```


<br>

# `xl()`

`xl()` is the primary function of the package 'xlr', and many of the
other functions in the package were developed to help it. The function
is somewhat similar to `View()` from the `utils` package in that `xl()`
is for quickly viewing datasets.

It's easiest to see what `xl()` does by passing it some example data;
try sending it some pre-loaded datasets, like:

```{r xlr-example1, eval = F}
xl(mtcars, iris, letters, AirPassengers, Titanic)
```

If you type that code to the R console, you should get a workbook that
pops open automatically with those datasets presented in standardized,
formatted sheets. Once you're done with the workbook, all you have to do
is close it; your spreadsheets are temporary. Close them, and they're
gone.

The video below is of `xl()` creating a temporary workbook where each
data object becomes a labeled spreadsheet in the workbook.

![xl(iris, mtcars, 1:100, dplyr::starwars)](www/Screencast-xl.webm)


To use `xl()`, pass a dataframe or lists of dataframes. It does not
matter if you pass in lists with nested lists where your data objects are
buried somewhere. `xl()` recursively flattens and auto-names the spreadsheets and 
workbook. In fact, nested lists can be useful in the naming of
groupings within spreadsheet tabs if you provide a sheet-naming specification 
like `.tabname_spec = list(name_spec = "{outer}|{inner}")`.

Try running the below examples in your console.

```
your_datasets <- enlist(flowers = iris, mtcars, datasets::USArrests)

xl(your_datasets) 

xl(your_datasets, .tabname_spec = list(name_spec = "{outer}|{inner}")) 

iris |> split(f = iris$Species) |> xl()

iris |> split(f = iris$Species) |> xl(your_datasets)

```


`xl()` writes your workbooks in a temp directory and, to keep workbooks
from building up, uses `later::later()` to remove the savefile 5 minutes
after write time. Temp directories also delete themselves when you
close R. If, however, you want to save a workbook, specify a path or
just a save name in the `xl()` input parameter `.path`. Specifying a
savename signals to `xl()` not to mark the file for cleanup. For
example, `xl(iris, .path = 'flower_data')` will save your data under
`flower_data.xlsx` within your working directory. (Note that parameter `.path`
uses `here::here()` internally.)


`xl()` creates tibbles of all datasets you pass in, and if specified by
the user parameter `.return`, optionally returns (always invisibly) the
`openxlsx` workbook object (default), or the data objects in a list, or
the workbook savepath as a string.

<br>

# `enlist()`

`enlist()` is for quick list-naming. This function produces named lists,
without list-nesting when passed a single bare list. That is, unlike
`list()`, `enlist()` removes its own embedded list layer if the user
passes in a single list.

For instance, compare the structure and naming created by enlist:

```{r enlist-example1, eval = T}
enlist(enlist(enlist('hello'))) |> str()
```

To the structure and naming created by list:

```{r enlist-example2, eval = T}
list(list(list('hello'))) |> str()
```

The `enlist()` example is one list deep, whereas the `list()
three lists deep. However, if you send two list objects into enlist, as
shown in the example below, enlist has no choice but to wrap them in a
new list.

```{r enlist-example3, eval = T}
enlist(list(letters), list(letters)) |> str()
```

### A few additional notes about `enlist()`:

-   Naming is permissive, you can have duplicate names in a list, so be
    careful about indexing by name since R returns only the first
    equivalently named observation from the list.

-   `enlist()` has input parameter `.label`, that takes a naming
    function. Internally, `.label` is passed to `nm` in
    `rlang::set_names()` and so takes the same inputs. For example:
    `enlist('hi','hello', .label = ~paste0(.,'_',1:2))`.

-   `enlist()` supports non-standard evaluation, e.g. `enlist(!!!letters)`.

-   Ignores any empty inputs, e.g. `enlist(,,'hi',,)`.

### Some clipboard prep notes:

The next 3 `xlr` functions, use the clipboard memory. To use the
clipboard, install the `clipr` package and then check that the package
works by running `clipr::dr_clipr().` Also, if using Linux, like me,
make sure to install a clipboard tool via your terminal, e.g.
`sudo apt-get install xclip`.

<br>

# `copy_for_xl()`

`copy_for_xl()` copies data to your clipboard memory that's formatted
for pasting to cells in a spreadsheet. You can pass a data object to it,
like:

```{r copy_for_xl-example, eval = F}
copy_for_xl(mtcars)
```

### The `copy_for_xl()` key-chord:

However, a copying method that is way, way better is to use the quick
keys `ctrl + alt + shift + c`. To use these quick keys, run the
function, `set_xlr_key_chords()`. (To set up quick keys, you only ever
have to run this function once in RStudio, but you have to close down
RStudio and the re-open for the new quick keys to be active.) Then:

1)  Type `mtcars` in an RStudio text editor window (not in the console
    window)

2)  Use your cursor to highlight the word `mtcars`, and

3)  Press `ctrl + alt + shift + c` all together. (It's just like copying
    with `ctrl + c`, but instead use all 3 control keys, plus the `c`.)

You should see `✔ Table copied to clipboard:` written in the console.
Open a spreadsheet and paste the copied `mtcars` data to it.

The below video example uses `copy_for_xl()`, via the console, to copy
some starwars data.
![copy_for_xl(dplyr::starwars)](www/Screencast-copy_for_xl.webm)

<br>

# `paste_from_xl()`

`paste_from_xl()` pastes data from your clipboard memory to an R tibble
object. You can pass spreadsheet data to R by copying data from your
spreadsheet, and then by typing the following into the console:

```{r paste_from_xl-example, eval = F}
my_data <- paste_from_xl()
```

The function guesses whether you have field names present, but if the
function guessed wrong, you can specify with the parameter
`has_fieldnames`, like below:

```{r paste_from_xl-example2, eval = F}
my_data <- paste_from_xl(has_fieldnames = TRUE)
```

### The `paste_from_xl()` key-chord:

However, a pasting method that is way, way better is to use the quick
keys `ctrl + alt + shift + v`. To use this key-chord, run the function,
`set_xlr_key_chords()` if you haven't already done so. Then:

1)  Copy some data from a spreadsheet.

2)  Click over to the RStudio console or to an RStudio editor. (The
    console and editor have different paste_from_xl() functionality.)

3)  Press `ctrl + alt + shift + v` all together to paste. (Just like
    pasting with `ctrl + v`, but using all 3 control keys, plus the
    `v`.)

You should see a tibble of your spreadsheet data assigned to a variable
and the dataset echoed to the console. However, if you pressed
`ctrl + alt + shift + v`, and you were over an editor window instead of
the console, you also get formatted code written into the editor window
that recreates the data object from script. The script is a formatted
version of R's deparsed internal data representation.

The video below is an example of copying spreadsheet data and then using
key-chord `ctrl + alt + shift + v` in RStudio.

![ctrl + alt + shift + v](www/Screencast-paste_from_xl.webm)

### Copy paths using `paste_from_xl()`:

In addition to spreadsheet data, you can copy file paths with
`paste_from_xl()`. Open a file navigation window in your operating
system, highlight a few files or folders, and press `ctrl + c`; then go
back to RStudio, click on an editor window, and press the key-chord
`ctrl + alt + shift + v`, you get an enscripted vector of file paths
pasted into your text editor.

<br>

The video below is an example of copying file paths from a file
window and then using the key-chord `ctrl + alt + shift + v` in RStudio.

![ctrl + alt + shift + v](www/Screencast-paste_from_xl-paths.webm)

<br>

# `enscript()`

I often manually process data down to particular information that's too
small to deal with saving and retrieving from a serialized
file or database. Instead, I just want to save my processed dataset as text
within the script, but I hate manually recreating the object.

If you have an object in a variable, and want that variable's data to be
deparsed into a formatted script, pass that variable into `enscript()`,
as in:

```{r enscript-example1, eval = F}
enscript(head(mtcars))
```

`enscript()` copies formatted code to your clipboard that will reproduce
that variable's internal R object. 

Below is an example video of `enscript()`:
![enscript(enlist('hi', 'bye', head(iris,3)))](www/Screencast-enscript.webm)


Try running something like:
```{r enscript-example2, eval = F}
iris |> split(f = iris$Species) |> purrr::map(head, 2) |> enscript()
```

You can pass in an expression or a single variable by piping 'enscript()' to the end of 
anything, or by using the quick keys. Enscript(), and its key-chord, is extremely 
useful for creating unit tests for package test files.


### The `enscript()` key-chord:

It's convenient to use quick keys with `enscript()`, which are
`ctrl + alt + shift + n`. (I remember the key-chord as all three control
keys, plus the `n`, which sounds short for 'enscript'.)

To use the `enscript()` quick keys, Run the quick-key setup with
`set_xlr_key_chords()` if you haven't already. Then:

1)  Type a variable name like `mtcars` in your RStudio editor
    window (not in the console).

2)  Use your cursor to highlight the word `mtcars`.

3)  Press `ctrl + alt + shift + n` all together.

You should see the success message, `✔ Script copied to clipboard:`,
written to your console with the deparsed object echoed below it. You can
now paste a formatted script into an editor window or the console with
`ctrl + v`.

Note that the `enscript()` quick keys only work for a highlighted
variable or highlighted expression in the RStudio editor window; the
quick keys do not interpret anything highlighted in the console.

<br>

# `repeated()`

`repeated()` helps filter a dataframe for identical observations. It should 
probably be in `dplyr`, but it's in `xlr` because it's not in `dplyr` and I use it 
so often with `xl()` to understanding and sharing duplication errors. Use it like this:

```{r repeated-example, eval = T}
mtcars |> dplyr::filter(repeated(disp))
```

`repeated()` is just `x %in% x[duplicated(x)]` with assertion
checks, but way easier to remember than `x %in% x[duplicated(x)]`. Use
it in `and` and `or` statements like this:

```{r repeated-example2, eval = F}
# AND statements
mtcars |> dplyr::filter(repeated(hp), repeated(disp))
# OR statements
mtcars |> dplyr::filter(repeated(hp) | repeated(disp))
```

<br>

# `sys_open()`

`sys_open()` opens the .xlsx files, but also opens any file, folder, or URL
since this function requests the operating system to open the provided input
locations via your operating system's command-line interface. This function
passes a path string (or many, if you like) to your OS, asking the OS to open up
the location using its default application. The reason I use it instead of other
functions is because I've had better luck getting files to open when requesting 
that the operating system make the decisions about how to open up the file.

If you aren't able to open a file with `sys_open()`, then, using your
operating system's navigation window, right-click on the file to make
sure your operating system has a default application assigned that file
type; after that, double-click on the file to make sure it opens in your
specified default application.

The video below is an example of opening up some PDF files by piping
them from `paste_from_xl()` to `sys_open()`.

![sys_open(paste_from_xl())](www/Screencast-sys_open.webm)

<br>

# `list_iron()`

`list_iron()` takes any nested list of lists and irons the input out to
get a single-file list of objects.

Because an '.xlsx' workbook can only make spreadsheets out of a simple,
flat list of data objects, `list_iron()` recursively smooshes any
additional list structure out of the input. `list_iron()` also has a few
convenience features for naming the list objects and applying a function
to each. For example, consider the ridiculously embedded list structure
below and list_iron's simplification of it:

```{r list_iron-example, eval = T}
ridiculous <- list(list(car_data = mtcars, list(list(flower_data = iris))))

# Original list depth
ridiculous |> str()

# Ironed list depth
ridiculous |> list_iron() |> str()
```

Notice the auto-naming of the list items above, i.e.
*"ridiculous|car_data"*, *"ridiculous|flower_data"*. Control the naming with
`name_spec` and `name_repair`. You can also apply a function, with parameter `.f`,
while flattening out all of the leaves in the lists , as shown below:

```{r list_iron-example2, eval = T}
list_iron(ridiculous, .f = ~tail(.,2))
```

<br>

# `entibble()`

`entibble()` makes a tibble with defaults useful for spreadsheets. Defaults include: 

* automatically presenting rownames, if they exist, as the first column of the tibble
* not repeating input elements to obtain common lengths when producing a dataframe
* treating a list of data inputs equivalent to comma-separated data inputs

For instance, compare the tibbled and entibbled data below.

```{r entibble-example, eval = T}
alpha <- rlang::set_names(letters, LETTERS)

# rownames are forced in so they show up in spreadsheets
entibble(alpha) |> head(2)
tibble::tibble(alpha) |> head(2)

# lists passed to entibble are dumped and their objects joined like comma-separated 
# inputs if the listed objects have compatible row dimensions
entibble(list(alpha, letters)) |> tail(2)
tibble::tibble(list(alpha, letters)) |> tail(2)

```
